---
title: "Covariate Extraction"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
always_allow_html: true
---

```{r  setup, include=FALSE}
# Set seed for reproducibility
set.seed(42) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
```

```{css, echo = FALSE, include = FALSE}
body .main-container {
  max-width: 100% !important;
  width: 100% !important;
}
body {
  max-width: 100% !important;
}

body, td {
  font-size: 16px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
.custom-box {
  background-color: #f5f7fa; /* Light grey-blue background */
    border-color: #e1e8ed; /* Light border color */
    color: #2c3e50; /* Dark text color */
    padding: 15px; /* Padding inside the box */
    border-radius: 5px; /* Rounded corners */
    margin-bottom: 20px; /* Spacing below the box */
}
.caption {
  margin: auto;
  text-align: center;
  margin-bottom: 20px; /* Spacing below the box */
}
```

In this section, we will show how to extract the spatial covariates from the OpenStreetMap data. We will compute the distances to the nearest objects and apply transformations to them. We will use the Euclidean distance to the nearest mosque, school, bank, fuel station, and hospital. We will then apply different transformations to these distances, such as the inverse, inverse square root, logarithm, and exponential decay. 

# Extracting the Spatial Covariates
We will now create `additional_city_objects.RData' file that contains the spatial covariates. We will extract the following spatial covariates: mosques, education places, hospital places, finance places, and intersections. 

```{r}
library(osmdata)
library(sf)
library(MetricGraph)
library(rSPDE)
library(INLA)
library(qs)
library(ggplot2)
library(leaflet)
library(dplyr)
library(here)

# Define the CRS using the WKT string
crs_wkt <- 'GEOGCRS["unknown",
    DATUM["World Geodetic System 1984",
        ELLIPSOID["WGS 84",6378137,298.257223563,
            LENGTHUNIT["metre",1]],
        ID["EPSG",6326]],
    PRIMEM["Greenwich",0,
        ANGLEUNIT["degree",0.0174532925199433],
        ID["EPSG",8901]],
    CS[ellipsoidal,2],
        AXIS["longitude",east,
            ORDER[1],
            ANGLEUNIT["degree",0.0174532925199433,
                ID["EPSG",9122]]],
        AXIS["latitude",north,
            ORDER[2],
            ANGLEUNIT["degree",0.0174532925199433,
                ID["EPSG",9122]]]]'

# Create the CRS object
new_crs <- st_crs(crs_wkt)

# Read the polygon data and set the new CRS
polygon <- st_read(here("Data/center_ahsa_polygon.geojson")) %>%
  st_cast("POLYGON") %>%
  st_set_crs(new_crs)
bb <- st_bbox(polygon)

# Visualize the polygon boundary
leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = polygon, color = "red", fill = FALSE, weight = 2)

# Load graph
graph <- qread(here("Data/graph_mesh_pruned_100m_7FRC_ver2.qs"))
```

## Highways and Intersections
```{r}

####### EXPLORING available OSM Data #######
data_width = opq(bb = bb) %>%
  add_osm_feature(key = "width") %>%
  osmdata_sf() 

## HIGHWAY
tags_highway <- available_tags("highway")
data_highway <- opq(bb = bb) %>%
  add_osm_feature(key = "highway", value = tags_highway$Value) %>%
  osmdata_sf()

points_highway <- data_highway$osm_points %>%
  filter(!is.na(highway)) %>%
  select(geometry, highway)
table(points_highway$highway)
graph$plot(vertex_size = 0, edge_width = 0.1) +
  geom_sf(data = points_highway, aes(color = highway)) +
  theme_minimal()

# discard street_lamp
points_highway <- points_highway %>% filter(highway != "street_lamp")

lines_highway <- data_highway$osm_lines %>%
  filter(!is.na(highway)) %>%
  select(geometry, highway)
table(lines_highway$highway)

polygon_highway <- data_highway$osm_polygons %>%
  filter(!is.na(highway)) %>%
  select(geometry, highway)
table(polygon_highway$highway)

graph$plot(vertex_size = 0, edge_width = 0.1) +
  geom_sf(data = polygon_highway, aes(color = highway)) +
  theme_minimal()

```

## Leisure, Amenities, and Buildings
```{r}
## LEISURE
tags_leisure <- available_tags("leisure")
data_leisure <- opq(bb = bb) %>%
  add_osm_feature(key = "leisure", value = tags_leisure$Value) %>%
  osmdata_sf()
points_leisure <- data_leisure$osm_points %>%
  filter(!is.na(leisure)) %>%
  select(geometry, leisure)
table(points_leisure$leisure)

polygon_leisure <- data_leisure$osm_polygons %>%
  filter(!is.na(leisure)) %>%
  select(geometry, leisure)
table(polygon_leisure$leisure)


## AMENITY
tags_amenity <- available_tags("amenity")
data_amenity <- opq(bb = bb) %>%
  add_osm_feature(key = "amenity", value = tags_amenity$Value) %>%
  osmdata_sf()
points_amenity <- data_amenity$osm_points %>%
  filter(!is.na(amenity)) %>%
  select(geometry, amenity)
table(points_amenity$amenity)[order(table(points_amenity$amenity), decreasing = TRUE)]

## BUILDING
tags_building <- available_tags("building")
data_building <- opq(bb = bb) %>%
  add_osm_feature(key = "building") %>%
  osmdata_sf()

points_building <- data_building$osm_points %>%
  filter(!is.na(building)) %>%
  select(geometry, building)
table(points_building$building)[order(table(points_building$building), decreasing = TRUE)]

polygon_building <- data_building$osm_polygons %>%
  filter(!is.na(building)) %>%
  select(geometry, building)
table(polygon_building$building)[order(table(polygon_building$building), decreasing = TRUE)]


```

## Main Functions
```{r}
######### EXTRACTING Specific Amenities #########
remove_duplicates_within_threshold <- function(data, threshold) {
    # Convert the distance threshold to meters
    distance_threshold <- units::set_units(threshold, "meters")

    # Identify pairs of points that are within the distance threshold
    within_distance <- st_is_within_distance(data, data, dist = distance_threshold)
    keep <- rep(TRUE, length(within_distance))

    # Loop through each point and mark duplicates for removal
    for (i in seq_along(within_distance)) {
      if (keep[i]) {
        duplicates <- within_distance[[i]]
        # Mark all duplicates (excluding the current point) for removal
        if (length(duplicates) > 1) {
          keep[duplicates[-1]] <- FALSE
        }
      }
    }
    unique_data <- data[keep, ]
    return(unique_data)
  }


# Fetch OSM data for both points and polygons
fetch_osm_data <- function(key, value, bb) {
  # Fetch points data and check if it exists
  data_points <- tryCatch({
    opq(bb = bb) %>%
      add_osm_feature(key = key, value = value) %>%
      osmdata_sf() %>%
      .$osm_points %>%
      filter(!is.na(!!sym(key))) %>%
      select(geometry, !!sym(key))
  }, error = function(e) {
    NULL  # Return NULL if points data is not available
  })

  # Fetch polygons data and check if it exists
  data_polygons <- tryCatch({
    opq(bb = bb) %>%
      add_osm_feature(key = key, value = value) %>%
      osmdata_sf() %>%
      .$osm_polygons %>%
      filter(!is.na(!!sym(key))) %>%
      select(geometry, !!sym(key))
  }, error = function(e) {
    NULL  # Return NULL if polygons data is not available
  })

  # Check if both points and polygons data are available
  if (is.null(data_points) && is.null(data_polygons)) {
    return(NULL)  # Skip if no data available
  }

  # Convert polygons to points (centroids) if polygons are available
  if (!is.null(data_polygons) && nrow(data_polygons) > 0) {
    centroids_polygons <- st_centroid(data_polygons) %>%
      select(geometry, !!sym(key))
  } else {
    centroids_polygons <- NULL
  }

  # Combine points and centroids if both are available
  if (!is.null(data_points) && nrow(data_points) > 0 && !is.null(centroids_polygons)) {
    combined <- rbind(data_points, centroids_polygons)
  } else if (!is.null(data_points) && nrow(data_points) > 0) {
    combined <- data_points
  } else {
    combined <- centroids_polygons
  }

  # Check that only the value of interest is present
  combined <- combined %>% filter(!!sym(key) == value)

  unique_combined = remove_duplicates_within_threshold(combined, 100)

  return(unique_combined)
}

```

## Extracting the Amenities
```{r}
# MOSQUES  
mosques = fetch_osm_data("amenity", "place_of_worship", bb)
mosques2 = fetch_osm_data("building", "mosque", bb)

mosques2 = mosques2 %>% rename(amenity = building)
mosques = rbind(mosques, mosques2)
mosques$amenity = "mosque"
mosques = remove_duplicates_within_threshold(mosques, 200)
cat("The number of unique mosques is", nrow(mosques), "\n")
graph$plot(vertex_size = 0, edge_width = 0.1) +
    geom_sf(data = mosques, aes(color = amenity)) +
    theme_minimal()

# EDUCATION
college = fetch_osm_data("amenity", "college", bb)
college2 = fetch_osm_data("building", "college", bb)
school = fetch_osm_data("amenity", "school", bb)
school2 = fetch_osm_data("building", "school", bb)
kindergarten = fetch_osm_data("amenity", "kindergarten", bb)
library = fetch_osm_data("amenity", "library", bb)
university = fetch_osm_data("amenity", "university", bb)
university2 = fetch_osm_data("building", "university", bb)

# Merge all the education data and check the number of unique points
education = rbind(college, school, kindergarten, library, university)
education2 = rbind(college2, school2, university2)%>% rename(amenity = building)
education = rbind(education, education2)
education$amenity = "education"
education_places = remove_duplicates_within_threshold(education, 200)
cat("The number of unique education points is", nrow(education_places), "\n")
graph$plot(vertex_size = 0, edge_width = 0.1) +
    geom_sf(data = education_places, aes(color = amenity)) +
    theme_minimal()

# HEALTH
hospital = fetch_osm_data("amenity", "hospital", bb)
hospital2 = fetch_osm_data("building", "hospital", bb)
clinic = fetch_osm_data("amenity", "clinic", bb)
dentist = fetch_osm_data("amenity", "dentist", bb)
doctors = fetch_osm_data("amenity", "doctors", bb)
pharmacy = fetch_osm_data("amenity", "pharmacy", bb)

hospitals = rbind(hospital, clinic, dentist, doctors, pharmacy)
hospitals2 = hospital2 %>% rename(amenity = building)
hospitals= rbind(hospitals, hospitals2)
hospitals$amenity = "hospital"
hospital_places = remove_duplicates_within_threshold(hospitals, 200)
cat("The number of unique hospital points is", nrow(hospitals), "\n")
graph$plot(vertex_size = 0, edge_width = 0.1) +
    geom_sf(data = hospital_places, aes(color = amenity)) +
    theme_minimal()


# FINANCE
bank = fetch_osm_data("amenity", "bank", bb)
atm = fetch_osm_data("amenity", "atm", bb)

finance = rbind(bank, atm)
finance$amenity = "finance"
finance_places = remove_duplicates_within_threshold(finance, 200)
cat("The number of unique finance points is", nrow(finance_places), "\n")
graph$plot(vertex_size = 0, edge_width = 0.1) +
    geom_sf(data = finance_places, aes(color = amenity)) +
    theme_minimal()

# JUNCTIONS
intersection = remove_duplicates_within_threshold(points_highway, 300)
# rename the column 
intersection = intersection %>% rename(amenity = highway)
intersection$amenity = "intersection"
graph$plot(vertex_size = 0, edge_width = 0.1) +
  geom_sf(data = intersection, aes(color = amenity)) +
  theme_minimal()

## Save all amenities data
save(mosques, education_places, hospital_places, finance_places, intersection, file = "Data/additional_city_objects.RData")

```


  
```{r}
# Plot all amenities
graph$plot(vertex_size = 0, edge_width = 0.1) +
  geom_sf(data = mosques, aes(color = amenity), size = 0.5) +
  geom_sf(data = education_places, aes(color = amenity), size = 0.5) +
  geom_sf(data = hospital_places, aes(color = amenity), size = 0.5) +
  geom_sf(data = finance_places, aes(color = amenity), size = 0.5) +
  geom_sf(data = intersection, aes(color = amenity), size = 0.5) +
  theme_minimal()
```





